### 2009-12-10 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Reviews of chapter 4 of the report... Adding a label on chapter 2 and adding references to non-cited sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@689 559edde2-a4d4-44e0-96fc-4035e40d7396
Revision on the feedback, chapter 2. -	Fixing typos -	Changing from temperature-scale to temperature unit -	Giving a practical example of transaction-time -   Giving a practical example about specific data location (precision) -   No additional information added for radar data produced (no references)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@691 559edde

\section{Persistence Storage for Sensor Networks}

@@

In general, sensor devices produce and preserve data onto their local
storage or transmit them to data nodes as described earlier in this chapter. It
is vital to understand that the majority of these devices output the raw data
in streams of values with any type of delimitation between each other. For
this reason, the processes of parsing and labeling each of the values located
in the data stream regularly requires the use of the device's specification
documentation. Similarly, in order to have access to each of the properties of
collected data by a given property related to time or space,
\cite{sn-provenance} also mentions that Metadata\footnote{Data about data}
should be included together with the collected data.

@@

A different way to describe sensor devices is through the use of
Annotations. \cite{sn-annotation} shows the use of tags to describe data
collected from a physical area monitoring. The sensor network was
designed to monitor moving objects, capturing their location while in
movement and their visual appearance using a video camera. In this way,
adding tags to particular video frames can describe different them, and
therefore, providing an efficient way to search for a specific frame based on
its visual observations. Similarly, the above referenced project also uses data
provenance to describe its collected data, since each of the collected data 
contains the related GPS coordinates of the observed object.

@@

Briefly, the collected data is described by the name of the observation
property. For instance, \textbf{temperature} = 73.45 expresses the value of the
observation of the temperature. In other words, the string ``temperature'' is
the property describing the raw collected data 73.45. On the other hand,
the example \textbf{temperature-scale} = ``Fahrenheit" describes the metadata
regarding the temperature.

@@

Based on the guidelines defined by \cite{db-temporal}, the time properties of
the collected data can be divided into two different dimensions: \textbf{valid
time} and the \textbf{transaction time}. The former depicts the point in time
when the observation took place with respect to the real-world space. An
important aspect of this metadata is that since it captures the time when the
observation occurred, its value cannot be changed during its existence, as well
as placing restrictions on the update operations of this constant. Similarly,
the latter type of is the point in time at which a given fact was transferred
to the persistence layer. Either types of time can be used depending on the
requirements of the sensor network. For instance, \cite{sn-dataware-house}
reports that all the data stored by its sensor network uses the time of
collection, but do not use the transaction time for that purpose.

@@

Depending on the application of the collected sensor data, the location in
which the data was collected may be needed. For example, \cite{sn-geo-metadata}
shows the geographic location of the collected data is critical to identify
the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
the position of any event captured by a black-box of an aircraft carries 
metadata for the event. Examples of this are the metadata event =
``high-pressure" of 1304, at the altitude of 1,500 feet above the ocean on GPS
coordinates of latitude 13.445 and longitude -23,003. This example clearly
gives necessary identification of where the event of high-pressure occurred.

>>>>

\section{Persistence Storage for Sensor Networks}
\label{sec:sn-persitence-storage}

@@

In general, sensor devices produce and preserve data onto their local
storage or transmit them to data nodes as described earlier in this chapter. It
is vital to understand that the majority of these devices output the raw data
in streams of values with any type of delimitation between each other. For
this reason, the process of parsing and labeling each of the values located
in the data stream regularly requires the use of the device's specification
documentation. Similarly, in order to have access to each of the properties of
collected data by a given property related to time or space,
\cite{sn-provenance} also mentions that Metadata\footnote{Data about data}
should be included together with the collected data.

@@

A different way to describe sensor devices is through the use of
Annotations. \cite{sn-annotation} shows the use of tags to describe data
collected from a physical area monitoring. The sensor network was
designed to monitor moving objects, capturing their location while in
movement and their visual appearance using a video camera. In this way,
adding tags to particular video frames can describe each of them, and
therefore, providing an efficient way to search for a specific frame based on
its visual observations. Similarly, the above referenced project also uses data
provenance to describe its collected data, since each of the collected data 
contains the related GPS coordinates of the observed object.

@@

Briefly, the collected data is described by the name of the observation
property. For instance, \textbf{temperature} = 73.45 expresses the value of the
observation of the temperature. In other words, the string ``temperature'' is
the property describing the raw collected data 73.45. On the other hand,
the example \textbf{temperature-unit} = ``Fahrenheit" describes the metadata
regarding the temperature.

@@

Based on the guidelines defined by \cite{db-temporal}, the time properties of
the collected data can be divided into two different dimensions: \textbf{valid
time} and the \textbf{transaction time}. The former depicts the point in time
when the observation took place with respect to the real-world space. An
important aspect of this metadata is that since it captures the time when the
observation occurred, its value cannot be changed during its existence, as well
as placing restrictions on the update operations of this constant. Similarly,
the latter type of is the point in time at which a given fact was transferred
to the persistence layer. Either types of time can be used depending on the
requirements of the sensor network. For instance, \cite{sn-dataware-house}
reports that all the data stored by its sensor network uses the time of
collection, but do not use the transaction time for that purpose. Similarly,
\cite{sn-provenance} shows a practical examples of critical queries performed
over the transaction time, when doctors need all the collected data regarding
the heart beat rate from the moment the patient entered the ambulance until
the moment of arrival at the hospital.

@@

Depending on the application of the collected sensor data, the location in
which the data was collected may be needed. For example, \cite{sn-geo-metadata}
shows the geographic location of the collected data is critical to identify
the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
the position of any event captured by a black-box of an aircraft carries 
metadata for the event. Examples of this are the metadata event =
``high-pressure" of 1304, at the altitude of 1,500 feet above the ocean on GPS
coordinates of latitude 13.445 and longitude -23,003. This example clearly
gives necessary identification of where the event of high-pressure occurred.
The precision of the location can be important for the collected data because
it helps researchers correlate different measurements based on exact locations
of the data. For instance, \ref{sn-provenance} depicts an example where two
different sensor networks covering the same physical area: one covering the
reginal traffic and another snowfall. Although each of them has individual
functions, collectively they could allow real-time navigation service to answer
queries regarding expected traffic patterns based on the current snowfall and
advise drivers accordingly by reusing collected data from both sensor networks.
